An Open Repository of Real-Time COVID-19 Indicators

### 2021-06-14 indicators/paper/paper.tex
Subsections on viz and forecasting

% RJT: @AR I'm now thinking we should add a very short paragraph or two on
% this. We can say that we provide various interactive maps highlighting both
% temporal (time series) and spatial aspects (choropleth or beehive maps), but
% also exploratory tools for correlations and anomalies.  (Very much like the
% correlations computed in the next section.)

% Another reason why I'm in favor of this is that we can now add everybody as a
% middle author who now contributed anything at all to COVIDcast, from database
% to API to indicator to visualization.  I'm sure the HCII students & faculty
% who worked with us would very much appreciate being on this paper.

@@

The indicators that are available in the COVIDcast API have been used in
dashboards produced by COVID Act Now \cite{CovidActNow}, COVID Exit Strategy
\cite{CovidExitStrategy}, and others; to inform the Delphi, DeepCOVID
\cite{Rodriguez:2021}, and the Institute for Health Metrics and Evaluation
(IHME) \cite{IHMEProj} COVID forecasting models; in various federal and state
government reports and analyses; and in a range of news stories. Aside from
operational use in decision-making and forecasting, they have also facilitated
numerous analyses studying the impacts of COVID-19 on the public, the
effectiveness of policy interventions, and factors that influenced the spread of
the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
  Doerr:2021}. The API currently serves hundreds of thousands of requests to
thousands of users, each day.

In what follows, we present illustrative examples of the usefulness of some of
the novel signals available in the API. These examples demonstrate that such
indicators are meaningfully related to COVID activity, that they provide
alternate views on pandemic activity that are not subject to the same reporting
glitches and delays as traditional public health surveillance streams, and that
they provide information about public behavior and attitudes that are not
available from any other source. Code to reproduce all examples (which uses the
\texttt{covidcast} R package and fetches data directly from the API) can be
found at
\url{https://www.github.com/cmu-delphi/covidcast-pnas/indicators/code/}.

@@

Many of the indicators in the COVIDcast API are intended to track COVID
activity. Five indicators in particular have the closest connections to
confirmed COVID cases:
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community):
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Quidel COVID test positivity rate (Quidel-TPR): The percentage of positive
  test results among Quidel COVID antigen tests.
\end{itemize}
Figure~\ref{fig:time_trends_national} compares the first three of these signals
to reported COVID cases in the United States (from JHU CSSE, and smoothed with a
7-day trailing average) over the first year of the pandemic (April 15, 2020 to
April 15, 2021), illustrating how they  appear to track national trends in cases
quite well. Importantly, this same relationship persists broadly across multiple
resolutions, down to smaller geographic regions such as states and counties, as
shown in the supplement. This will also be illustrated in a more detailed
correlation analysis in the next subsection.

@@

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations of
these five signals over the same time period, and for the same set of
counties. For each signal, we display the set correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
There are two remarks deserving attention:

@@

\begin{itemize}
\item This is different from what is observed in
  Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
  achieves clearly the highest correlations for most of the time
  period. However, it is worth emphasizing that time-wise and geo-wise
  correlations are truly measuring different properties of a signal; and the
  claims signals (CHNG-COVID and CHNG-CLI) seem more appropriate for
  temporal---rather than spatial---comparisons.  We revisit this point in the
  discussion.
\item It is still quite impressive and surprising that the CTIS-CLI-in-community
  signal, based on people reporting on the symptoms of others around them, can
  achieve  nearly as strong time-wise correlations to confirmed cases as can a
  signal that is based on picking up the occurrence of a confirmed case passing
  through the outpatient system.
\end{itemize}

>>>>

Since April 2020, Delphi has been maintaining and continually improving various
online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
These tools fetch data directly from the API, and allow for exploration of both 
temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
in the signals, as well as many other aspects, such as correlations, anomalies,
and backfill.  There is also a dedicated dashboard for exploring results from
the COVID Tracking and Impact Survey.

\subsection{COVID-19 Forecasting Systems}

Since July 2020, Delphi has been regularly submitting short-term forecasts of
COVID-19 case and death incidence, at the state and county levels, to the
COVID-19 Forecast Hub \cite{ForecastHub}, under the name ``CMU-TimeSeries''.
Building, training, and deploying our forecasting models leverages much of the
infrastructure described in this paper (such as the COVIDcast API's \textit{as
  of} feature), and some of our forecasting models rely on auxiliary indicators
(such as survey-based and claims-based COVID-like illness indicators, described
below, which we have used in case forecasting).

@@

The indicators that are available in the COVIDcast API have been used in
dashboards produced by COVID Act Now \cite{CovidActNow}, COVID Exit Strategy
\cite{CovidExitStrategy}, and others; to inform the Delphi, DeepCOVID
\cite{Rodriguez:2021}, and the Institute for Health Metrics and Evaluation
(IHME) \cite{IHMEProj} COVID forecasting models; in various federal and state
government reports and analyses; and in a range of news stories. Aside from
operational use in decision-making and forecasting, they have also facilitated
numerous analyses studying the impacts of COVID-19 on the public, the
effectiveness of policy interventions, and factors that influenced the spread of
the pandemic \cite{Adjodah:2021, Pierri:2021, Jewell:2021, Chakrabarti:2020,
  Doerr:2021}. The API currently serves hundreds of thousands of requests to
thousands of users every day.

In what follows, we present examples of the usefulness of some of the novel
signals available in the API. These examples demonstrate that such indicators
are meaningfully related to COVID activity, that they provide alternate views on
pandemic activity that are not subject to the same reporting glitches and delays
as traditional public health surveillance streams, and that they provide
information about public behavior and attitudes that are not available from any
other source. Code to reproduce all examples (which uses the \texttt{covidcast}
R package and fetches data directly from the API) can be found at
\url{https://www.github.com/cmu-delphi/covidcast-pnas/indicators/code/}.

@@

Many of the indicators in the COVIDcast API are intended to track COVID
activity. Five indicators in particular have the closest connections to
confirmed cases:
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community):
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Quidel COVID test positivity rate (Quidel-TPR): The percentage of positive
  test results among Quidel COVID antigen tests.
\end{itemize}
Figure~\ref{fig:time_trends_national} compares the first three of these signals
to reported COVID cases in the United States (from JHU CSSE, and smoothed with a
7-day trailing average) over the first year of the pandemic (April 15, 2020 to
April 15, 2021), illustrating how they  appear to track national trends in cases
quite well. Importantly, this same relationship persists broadly across multiple
resolutions, down to smaller geographic regions such as states and counties, as
shown in the supplement. This will also be illustrated in a more detailed
correlation analysis in the next subsection.

@@

@@

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from 
these five signals over the same time period, and for the same set of
counties. For each signal, we display the set correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
There are two remarks deserving attention:

\begin{itemize}
\item This is different from what is observed in
  Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
  achieves clearly the highest correlations for most of the time
  period. However, it is worth emphasizing that time-wise and geo-wise
  correlations are truly measuring different properties of a signal; and the
  claims signals (CHNG-COVID and CHNG-CLI) seem more appropriate for
  temporal---rather than spatial---comparisons.  We revisit this point in the
  discussion.
\item It is still quite impressive (and surprising) that the
  CTIS-CLI-in-community signal, based on people reporting on the symptoms of
  others around them, can achieve  nearly as strong time-wise correlations to
  confirmed cases as can a signal that is based on picking up the occurrence of
  a confirmed case passing through the outpatient system.
\end{itemize}
